Correctness of Algebraic File Synchronization

### 2015-11-21 2015/paper.tex
Finalizing text II

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\wr\!\!\wr\,} % Independent

@@

@@

@@

@@

\section{Defining the Algebra}

% -----------.........-------------.............--------------..............

\begin{mydef}
% Unsure if this relation should be in the algebra.
For a set of $k$ sequences
$\works{A_1,A_2,\ldots,A_k}$ means that 
$A_1,A_2,\ldots,A_k$ work at the same time, that is,
% intended interpretation:
\[\exists \FS: A_1\FS\neq\fsbroken \wedge \cdots \wedge A_k\FS\neq\fsbroken\]
As the sequences form a set, their order is irrelevant.
% inference rules:
Also, we see that 
\end{mydef}

\begin{mydef}
$\worksc{A_1,A_2,\ldots,A_n}{B_1,B_2,\ldots,B_n}$ means that 
all of $A_1,A_2,\ldots,A_n$ work where all of $B_1,B_2,\ldots,B_n$ work.
TODO: $A$s $B$s are sets.
We assume that the arguments $A_i$ and $B_i$ commute among each other,
and that the relation is reflexive, that is, $\worksc{B}{B}$ is true for all sequences.
Its intended interpretation is
\begin{align*}
\forall \FS:&\\ 
&B_1\FS\neq\fsbroken \wedge \cdots \wedge B_n\FS\neq\fsbroken\\
&\Rightarrow\\
&A_1\FS\neq\fsbroken \wedge \cdots \wedge A_n\FS\neq\fsbroken
\end{align*}
\end{mydef}

% INFERENCE RULES

\begin{myrul}\label{r_ex_w}
\[\infer{B\eqext A}{\worksc{A}{B}}\]
\end{myrul}

These rules also apply if the second, conditional part of $\worksc{\ldots}{B_1,\ldots,B_n}$ is empty:

\begin{myrul}
\[\infer{\worksc{A_1,\ldots,A_{n-1},A_n}{B_1,\ldots,B_n}}{\worksc{A_1,\ldots,A_{n-1}}{B_1,\ldots,B_n}}\]
\end{myrul}

\begin{myrul}\label{r_w_combine}
\[\inferr{\worksc{A_1,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{C}{B_1,\ldots,B_n}}{\worksc{A_1,\ldots,A_n,C}{B_1,\ldots,B_n}}\]
\end{myrul}

\begin{myrul}\label{r_w_w}
\[\inferr{\worksc{A_1,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{C}{A_n}}{\worksc{A_1,\ldots,A_n,C}{B_1,\ldots,B_n}}\]
\end{myrul}

\begin{myrul}{??? - and vice versa}
\[\infer{\works{A}}{A\nequiv\cbrk}\]
\end{myrul}

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)\in A$) and $B$ ($\czw(p, V_W)\in B$)
then their input types must match ($X=Z$).
TODO: Is this an axiom?
\end{mylem}

\begin{myrul}\label{r_invmove}
\[ \infer{\worksc{S\cc A}{B}}{\worksc{A}{S^{-1}\cc B}} \]
To see the soundness of this rule...
Sequences are bijections on filesystems over type-equality.
TODO: $A,S,B$ need to be simple?
\end{myrul}

We say that the nodes $n$ and $m$ are \textbf{unrelated} if $n\not\descendant m$ and $n\not\ancestor m$
and $n\new m$.

We aim to use our algebra to define and implement algorithms for conflict detection
and ordering commands before we can apply them to all replicas.
Conflicts and ordering all happen between commands, therefore in our algebra
we will reason about sequences of commands, 
and do so based on a set of axioms about pairs of neighbouring commands.
We are interested to see which pairs of commands cause errors all the time,
and which can be simplified or reversed
in order to see under what circumstances commands conflict or can be reordered.

Pairs of commands in general will have the form
\[ \cxy(p,V_Y)\cc  \czw(q,V_W) \]
where $X,Y,Z,W\in\typeset$, $p,q\in\setp$, 
and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.

The following axioms describe multiple statements
that we accept as true in the algebra.
These statements are generated by substituting 
all possible values for
$X,Y,Z,W\in\typeset$; $p,q,\pp\in\setp$; $V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$
under the conditions noted in the axiom.

\begin{myax}\label{ax_separate_commute}
Commands on unrelated paths commute:
$\cxy(p,V_Y)\cc  \czw(q,V_W) \equiv \czw(q,V_W)\cc  \cxy(p,V_Y)$ where $p\not\descendant q$ and $q\not\descendant p$.
TODO $p\neq q$
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated paths do not break every filesystem:
$\cxy(p,V_Y)\cc  \czw(q,V_W) \nequiv \cbrk$ where $p\not\descendant q$ and $q\not\descendant p$.
TODO $p\neq q$
\end{myax}

% TODO Use 'node' instead of 'path'

\begin{myax}\label{ax_same_breaks}
Commands on the same path break every filesystem if incompatible:
$\cxy(p,V_Y)\cc  \czw(p,V_W) \equiv \cbrk$ where $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_same_emptyseq}
Commands on the same path simplify:
$\cxy(p,V_Y)\cc  \czw(p,V_W) \eqext \emptyseq$ where $Y=Z$ and $X=W=\fsregcommandchar{\empt}$ or $X=W=\fsregcommandchar{D}$.
\end{myax}

\begin{myax}\label{ax_same_singlec}
Commands on the same path simplify:
$\cxy(p,V_Y)\cc  \czw(p,V_W) \equiv \cxw(p,V_W)$ where $Y=Z$ and $XW\neq\cbb$ and $XW\neq\cdd$.
\end{myax}

@@

@@

@@

TODO $\reca$ is maximal

>>>>

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\wr\!\!\wr\,} % Independent commands, sequences
\newcommand{\unrel}{\indep} % Unrelated nodes

@@

% Insight: (Prof. Norman Ramsey)
% - updates where two orders are not equivalent are incompatible
% - we only have a set of updates and need to create an order

@@

% Definitions for command pairs and ordering
% ------------------------------------------

@@

Let us also define the following relations between nodes:

@@

\begin{mydef}{$n\unrel m$.}
We write $n\unrel m$, or ``$n$ and $m$ are unrelated'' if $n\not\descendant m$ and $n\not\ancestor m$
and $n\new m$.
\end{mydef}

% Command pairs - Axioms
% ----------------------

\section{Investigating command pairs}

Our aim is to derive information about the effects of sequences
of commands independently of that actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we arguad above, we can disregard the exact output value of commands apart from its type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

We methodically investigated of all possibilities using a computer program
to determine
which pairs of commands cause errors all the time,
which can be simplified to one or no commands, and which can be reversed
without any change in their overall effect.
% TODO: Footnote: URL
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.

Below we list a number of statements derived using this method.
Although they can be proven by investigating all relevant possibilities
in a filesystem, we refer to these in axioms because in a secondary algebraic model
that describes commands without reference to an underlying filesystem,
these would function as meta-rules generating statements we would accept as true.
See section (TODO) for more about such a model.

Pairs of commands in general have the form
\[ \cxy(n,V_Y)\cc  \czw(m,V_W) \]
where $X,Y,Z,W\in\typeset$, $n,m\in\setn$, 
and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.

\begin{myax}\label{ax_separate_commute}
Commands on unrelated nodes commute:
$\cxy(n,V_Y)\cc  \czw(m,V_W) \equiv \czw(m,V_W)\cc \cxy(n,V_Y)$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated nodes do not break every filesystem:
$\cxy(n,V_Y)\cc  \czw(m,V_W) \nequiv \cbrk$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_same_breaks}
Commands on the same node break every filesystem if their types are incompatible:
$\cxy(n,V_Y)\cc \czw(n,V_W) \equiv \cbrk$ where $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_same_emptyseq}
Commands on the same node simplify:
$\cxy(n,V_Y)\cc \czw(n,V_W) \eqext \emptyseq$ where $Y=Z$ and $X=W=\fsregcommandchar{\empt}$ or $X=W=\fsregcommandchar{D}$.
\end{myax}

\begin{myax}\label{ax_same_singlec}
Commands on the same node simplify:
$\cxy(n,V_Y)\cc  \czw(n,V_W) \equiv \cxw(n,V_W)$ where $Y=Z$ and $XW\neq\cbb$ and $XW\neq\cdd$.
\end{myax}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

\begin{mylem}\label{r_invmove}
\[ \infer{\worksc{S\cc A}{B}}{\worksc{A}{S^{-1}\cc B}} \]
To see the soundness of this rule...
Sequences are bijections on filesystems over type-equality.
TODO: $A,S,B$ need to be simple?
\end{mylem}

@@

% Introducing 'works'
% -------------------

So that we can formalise statements (TODO) needed to prove this,
let us introduce two relations that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

\begin{mydef}
% Unsure if this relation should be in the algebra.
For a set of $k$ sequences
$\works{A_1,A_2,\ldots,A_k}$ means that 
$A_1,A_2,\ldots,A_k$ work at the same time, that is,
% intended interpretation:
\[\exists \FS: A_1\FS\neq\fsbroken \wedge \cdots \wedge A_k\FS\neq\fsbroken\]
As the sequences form a set, their order is irrelevant.
% inference rules:
We also see that $\works{A} \Leftrightarrow A\nequiv \cbrk$. 
\end{mydef}

\begin{mydef}
For two sets of sequences, $\worksc{A_1,A_2,\ldots,A_k}{B_1,B_2,\ldots,B_l}$ means that 
all of $A_1,A_2,\ldots,A_k$ work where all of $B_1,B_2,\ldots,B_l$ work,
that is,
% intended interpretation:
\begin{align*}
\forall \FS:&\\ 
&B_1\FS\neq\fsbroken \wedge \cdots \wedge B_k\FS\neq\fsbroken\\
&\Rightarrow\\
&A_1\FS\neq\fsbroken \wedge \cdots \wedge A_l\FS\neq\fsbroken
\end{align*}
As above, the order of the sequences in the sets is not relevant.
% inference rules:
It is easy to see that 
% $\worksc{X}{X}$ is always true for any set of sequences $X$,
$B\eqext A \Rightarrow \worksc{A}{B}$
and that this relation can be chained:
$ \worksc{A_1,\ldots,A_k}{B_1,\ldots,B_l} \wedge \worksc{C_1,\ldots,C_j}{A_1} \Rightarrow \worksc{A_1,\ldots,A_k,C_1,\ldots,C_j}{B_1,\ldots,B_l} $
\end{mydef}

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxy(n, V_Y)\in A$) and $B$ ($\czw(n, V_W)\in B$)
then their input types must match ($X=Z$).
\end{mylem}
\begin{proof}
This is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
$\cxy(n, V_Y)$ or $\czw(n, V_W)$ would not break, and consequently $A$ and $B$ could not work on the same filesystem.
\end{proof}

@@

% TODO $\reca$ is maximal

% TODO Talk about setting up an algebra / calculus
